\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2022-09-27 sections/abstract.tex
Change: Incorporate comments from Gabriel and add refs

\begin{abstract}
  {
    \noindent
    Microbes tend to organize into communities consisting of hundreds of species involved in complex interactions with each other.
    16S ribosomal RNA (16S rRNA) amplicon profiling provides snapshots that reveal the phylogenies and abundance profiles of these microbial communities.
    These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of associations in these communities.
    The inference of networks from 16S data involves numerous steps, each requiring specific tools and parameter choices.
    The extent to which the steps in this workflow affect the final network is still unclear.
    In this study, we perform a meticulous analysis of each step of a pipeline that can convert 16S sequencing data into a network of microbial associations.
    Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network and estimate steps that contribute most significantly to the variance.
    We further determine the tools and parameters that generate robust co-occurrence networks and develop consensus network algorithms based on benchmarks with mock and synthetic datasets.
    The Microbial Co-occurrence Network Explorer or \acs{micone} (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}), follows these default tools and parameters and can help explore the outcome of these combinations of choices on the inferred networks.
    We envisage that this pipeline could be used for integrating multiple data sets, and for generating comparative analyses and consensus networks that can help understand microbial community assembly in different biomes.
  }
\end{abstract}

@@

>>>>

\begin{abstract}
  {
    \noindent
    Microbes commonly organize into communities consisting of hundreds of species involved in complex interactions with each other.
    16S ribosomal RNA (16S rRNA) amplicon profiling provides snapshots that reveal the phylogenies and abundance profiles of these microbial communities.
    These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of associations in these communities.
    However, the inference of networks from 16S data involves numerous steps, each requiring specific tools and parameter choices.
    Moreover, the extent to which these steps affect the final network is still unclear.
    In this study, we perform a meticulous analysis of each step of a pipeline that can convert 16S sequencing data into a network of microbial associations.
    Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network and estimate steps that contribute most significantly to the variance.
    We further determine the tools and parameters that generate robust co-occurrence networks and develop consensus network algorithms based on benchmarks with mock and synthetic datasets.
    The Microbial Co-occurrence Network Explorer or \acs{micone} (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}), follows these default tools and parameters and can help explore the outcome of these combinations of choices on the inferred networks.
    We envisage that this pipeline could be used for integrating multiple datasets, and for generating comparative analyses and consensus networks that can help understand microbial community assembly in different biomes.
  }
\end{abstract}

@@

% \doublespacing

### 2022-09-27 sections/discussion.tex
Change: Incorporate comments from Gabriel and add refs

  We envision that \ac{micone}, and its underlying tools and databases, will be increasingly useful for building large comparative analyses across studies.
  It allows for rapid, configurable, and reproducible inference of microbial networks and further for the formulation of hypotheses about the role of these interactions on community composition and stability.
  These comparative analyses will require coupled network analysis and visualization tools (such as \ac{mind}~\cite{huResourceComparisonIntegration2022}) and will require more systematic access to datasets, shared in accordance with FAIR standards~\cite{pachecoFAIRRepresentationsMicrobial2022}.

>>>>

  We envision that \ac{micone}, and its underlying tools and databases, will be increasingly useful for building large comparative analyses across studies.
  It allows for rapid, configurable, and reproducible inference of microbial networks and further for the formulation of hypotheses about the role of these interactions on community composition and stability.
  These comparative analyses will require coupled network analysis and visualization tools (such as \ac{mind}~\cite{huResourceComparisonIntegration2022}) and need systematic access to datasets, shared in accordance with FAIR standards~\cite{pachecoFAIRRepresentationsMicrobial2022}.

### 2022-09-27 sections/introduction.tex
Change: Incorporate comments from Gabriel and add refs

  Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{Ghoul2016,Thompson2017} [MORE REFS].
  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships [REFS].
  Understanding how these interconnections shape community structure and functionalities is a fundamental challenge in microbial ecology, and has applications in the study of microbial ecosystems across different biomes.
  With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods,  more information can be extracted from these microbial community samples than ever before.
  In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons (hereafter referred to as 16S data) of microbial communities, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016,Lloyd-Price2016}.
  These advances have led to large-scale data collection efforts involving terrestrial (\acl{emp}) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}. [3 REFS for each]

 This wealth of information has the potential to help us understand how communities assemble and operate. In particular, a powerful tool for translating microbiome composition data into knowledge is the construction of association (or co-occurrence, or correlation) networks, in which  microbial taxa are represented by nodes, and frequent co-occurrences (or negative correlations) across datasets are encoded in edges between nodes.
 While the relationship between directly measured interactions [REFS] and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets.

 The importance of these networks is two-fold: first, they can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, and the community response to environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, they can serve as a bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
 For example, multiple studies have shown the importance of specific microbial  associations in the healthy microbiome \cite{Lloyd-Price2016} and the rise of dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
 In the context of terrestrial bio-geochemistry, co-occurrence networks were shown to help understand microbiome assembly \cite{Fierer2017}, and  the response of microbial communities to environmental perturbations \cite{Jiao2019}.

>>>>

Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{Zhang2015,Thompson2017,Ghoul2016,HumanMicrobiomeProjectConsortium2012,royo-llonchCompendium530Metagenomeassembled2021,tedersooFungalBiogeographyGlobal2014}.
  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships~\cite{lima-mendezDeterminantsCommunityStructure2015a,zelezniakMetabolicDependenciesDrive2015,coyteUnderstandingCompetitionCooperation2019,DSouza2018}.
  Understanding how these interconnections shape community structure and functionalities is a fundamental challenge in microbial ecology, and has applications in the study of microbial ecosystems across different biomes.
  With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods,  more information can be extracted from these microbial community samples than ever before.
  In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons (hereafter referred to as 16S data) of microbial communities, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016,Lloyd-Price2016}.
  These advances have led to large-scale data collection efforts involving terrestrial~\cite{Thompson2017,gilbertMeetingReportTerabase2010,tedersooFungalBiogeographyGlobal2014}, marine~\cite{Zhang2015,lima-mendezDeterminantsCommunityStructure2015a,royo-llonchCompendium530Metagenomeassembled2021} and human-associated microbiota~\cite{HumanMicrobiomeProjectConsortium2012,proctorIntegrativeHumanMicrobiome2019,Lloyd-Price2016}.

 This wealth of information has the potential to help us understand how communities assemble and operate.
 In particular, a powerful tool for translating microbiome composition data into knowledge is the construction of association (or co-occurrence, or correlation) networks, in which  microbial taxa are represented by nodes, and frequent co-occurrences (or negative correlations) across datasets are encoded in edges between nodes.
 While the relationship between directly measured interactions~\cite{lubbeExometabolomicAnalysisCrossFeeding2017,Jian2020,Hsu2019} and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017,Rottjers2018}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets~\cite{faustMicrobialInteractionsNetworks2012a,maEarthMicrobialCooccurrence2020a}.

 The importance of these networks is two-fold: first, they can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, and the community response to environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, they can serve as a bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
 For example, multiple studies have shown the importance of specific microbial associations in the healthy microbiome \cite{Lloyd-Price2016,Wu2016,HumanMicrobiomeProjectConsortium2012} and the rise of dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
 In the context of terrestrial biogeochemistry, co-occurrence networks were shown to help understand microbiome assembly \cite{Fierer2017}, and  the response of microbial communities to environmental perturbations \cite{Jiao2019}.

### 2022-09-27 sections/methods.tex
Change: Incorporate comments from Gabriel and add refs

  Note that we do not use Pearson and Spearman methods in the p-value merging step to determine the consensus network, these methods are only used for demonstration and comparison.
  The combined p-values are used to threshold for significance right before the consensus algorithm is applied to the inferred networks.

  \subsubsection*{Consensus method}
  \vspace{-5mm}
  The consensus algorithm was designed to increase the precision (number of true positives) at the end of the network inference step.
  For this purpose, we developed two simple algorithms that combine the edges reported by the different network inference tools using a consensus based on a parameter $p$.
  The inputs to both the algorithms are the co-occurrence networks (association matrices) $\bar{N}^i$ (flattened version of $N^{i}$) generated by each method $i$ and the threshold parameter $p$.
  Here, the $\bar{N}^{i}$ each have the same set of nodes $q$ and only differ by the value of the association inferred between every pair.
  Networks which do not have a particular node, are updated such that the node is added as an isolated component.
  In this manner, $\bar{N}^{i}_j$ represents edge $j$ in network $i$.

@@

  The simple voting method returns the union of the networks when $0 \leq p \leq \frac{1}{w}$ and will return the intersection when $(w - 1) \leq p \leq 1$.
  In general, if $(n - 1) \leq p \leq \frac{n}{w}$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.

>>>>

  Note that we do not use Pearson and Spearman methods in the p-value merging step to determine the consensus network.
  These methods are only used for demonstration and comparison.
  The combined p-values are used to threshold for significance right before the consensus algorithm is applied to the inferred networks.

  \subsubsection*{Consensus method}
  \vspace{-5mm}
  The consensus algorithm was designed to increase the precision (number of true positives) at the end of the network inference step.
  For this purpose, we developed two simple algorithms that combine the edges reported by the different network inference tools.
  Both the algorithms make use of a user-defined parameter $p$ ($0 \leq p \leq 1$), in order to threshold the edges from the individual methods.
  The inputs to both the algorithms are the co-occurrence networks (association matrices) $\bar{N}^i$ (flattened version of $N^{i}$) generated by each method $i$ and the threshold parameter $p$.
  Here, the $\bar{N}^{i}$ each have the same set of nodes $q$ and only differ by the value of the association inferred between every pair.
  Networks which do not have a particular node, are updated such that the node is added as an isolated component.
  In this manner, $\bar{N}^{i}_j$ represents edge $j$ in network $i$.

@@

  The simple voting method returns the union of the networks when $0 \leq p \leq \frac{1}{w}$ and will return the intersection when $\frac{(w - 1)}{w} \leq p \leq 1$.
  In general, if $\frac{(n - 1)}{w} \leq p \leq \frac{n}{w}$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.

### 2022-09-27 sections/results.tex
Change: Incorporate comments from Gabriel and add refs

  Since no database performs better than others against mock datasets, the choice of which database to use could be driven by other reasons (see Discussion).
  One reason to choose a particular database could be the frequency of updates and the potential for future growth.
  Both \ac{gg}, due to its frequent use in the literature~\cite{Balvociute2017}, and \ac{ncbi}, due to its regular revision and maintenance, could be good choices for taxonomy assignment, with few differences observed between the two in our analysis  (Figure~\ref{fig:figure_s5}).
  In our default pipeline, we choose \ac{gg} as the default method.

>>>>

  Since no database performs better than others against mock datasets, the choice of which database to use could be driven by other reasons (see Supplementary discussion).
  One reason to choose a particular database could be the frequency of updates and the potential for future growth.
  Both \ac{gg}, due to its frequent use in the literature~\cite{Balvociute2017}, and \ac{ncbi}, due to its regular revision and maintenance, could be good choices for taxonomy assignment, with few differences observed between the two in our analysis  (Figure~\ref{fig:figure_s5}).
  In our default pipeline, we choose \ac{gg} as the default method.

### 2022-09-27 sections/supplementary.tex
Change: Incorporate comments from Gabriel and add refs

  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated via a numerical integration:
  \begin{equation*}
    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right)
  \end{equation*}

  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
  \begin{equation}
    \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
    \label{eqn:suppl_covariance-pvalues}
  \end{equation}
  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$

  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
  \begin{equation}
    \begin{aligned*}
        & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
    \end{aligned*}
    \label{eqn:suppl_pvalue-combined}
  \end{equation}

  The p-value merging and consensus method in \ac{micone} (refer Methods) uses Equation~\ref{eqn:suppl_covariance-pvalues} to estimate the covariance of the p-values and Equation~\ref{eqn:suppl_pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
  The combined p-values are used to threshold for significance the correlation-based networks during the consensus network step.

>>>>

  We first calculate $E[\Psi]$ and $Var[\Psi]$ under the null hypothesis that the data are drawn from a multivariate Gaussian with some covariance matrix.
  We then use these values to parametrize a $\chi^2$ distribution from which p-value corresponding to $\frac{\psi}{c}$ can be calculated.
  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated via a numerical approximation:
  \begin{equation*}
    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right)
  \end{equation*}

  The above formulation was improved by Kost and McDermott~\cite{kost_combining_2002} by further fitting a third-order polynomial to approximate the covariance
  \begin{equation}
    \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
    \label{eqn:suppl_covariance-pvalues}
  \end{equation}
  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$

  Using $E[\Psi]$ and $Var[\Psi]$ one then fits a $\chi^2$ distribution with the parameters $c$ and $f$.
  Note that since, in general, $f$ will not be an integer, this should be understood as a Gamma distribution with a shape parameter $f$, as mentioned by Brown~\cite{brown_400_1975}.
  Using this, one calculates the test $\psi$ and computes the p-value from the CDF of the $\chi^2$ distribution, given in Equation~\ref{eqn:suppl_pvalue-combined}.
  Therefore, the final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
  \begin{equation}
    \begin{aligned*}
        & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
    \end{aligned*}
    \label{eqn:suppl_pvalue-combined}
  \end{equation}

  The p-value merging and consensus method in \ac{micone} (see Methods) uses Equation~\ref{eqn:suppl_covariance-pvalues} to estimate the covariance of the p-values and Equation~\ref{eqn:suppl_pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
  The combined p-values are used to threshold for significance the correlation-based networks during the consensus network step.

### 2022-09-27 sections/tables_figures.tex
Change: Incorporate comments from Gabriel and add refs

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=0.74\linewidth]{figure1.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The workflow of the \ac{micone} pipeline}.
      The steps of the workflow can be broken down into five major groups: \textbf{(SP)} \textbf{S}equence \textbf{P}rocessig, \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU and ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
      Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
      The text along the arrows describes the data that is being passed from one step to another.
      The final output of the pipeline is the consensus network generated from the inferred co-occurrence networks.
      For details on each process and the data outputs, see Methods.
    }
    \label{fig:figure1}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure2.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are lower in abundance.}
      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best-performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
      The data used for the analysis in \textbf{(C, D)} were the mock datasets from mockrobiota~\cite{Bokulich2016}.
    }
    \label{fig:figure2}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure3.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments beyond the Order level.}
      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
      This result illustrates how the same sequences are assigned to different Genus under different databases.
      A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
      The number of assigned genera for each database are displayed at the top of each column.
      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
      The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
      However, we still recommend the \ac{gg} database and the Naive Bayes classifier as the defaults for the \ac{micone} pipeline.
      Due to the popularity of the \ac{gg} database, comparison against existing networks in literature will be easier since many published datasets make use of this database in their analyses.
      The datasets used for the analysis in \textbf{(C)} were the mock datasets from mockrobiota.
    }
    \label{fig:figure3}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure4.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
      \textbf{(A)} The nine different networks (excluding \ac{mldm}) generated by the different network inference methods are very dissimilar.
      The green links are positive associations and the orange links represent negative associations.
      A threshold of 0.3 was set for the correlation-based methods (\ac{sparcc}, Propr, Spearman and Pearson) and a threshold of 0.01 was set for the other direct association methods.
      The correlation-based methods in general produce graph with higher edge-densities.
      \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large proportion of common nodes involved in connections (33 out of 68).
      Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
      The data used in this analysis were the healthy stool samples from the fecal microbiome transplant dataset.
      \ac{mldm} is not shown in the comparisons because the algorithm failed to converge for most of the network combinations.
    }
    \label{fig:figure4}
  \end{figure}

@@

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure6.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
      \caption{
      \textbf{The choice of reference database contributes to the most variance in the networks}.
      \textbf{(A)} The percentage of variance in the networks contributed by the DC, CC (chimera checking), TA and OP steps of the pipeline (right) and the linear model used to calculate the variance (left), see the Methods section for details.
      The data used for the analysis were the healthy stool samples from the fecal microbiome transplant dataset~\cite{Kang2017}.
      A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the analysis.
      The taxonomy database contributes most to the variance between the networks (65.4\%) followed by the filtering of the counts matrix (26.8\%) in the OP step.
      The variation due to the DC and CC steps are comparatively not as significant.
      The fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time, however, this can be ignored as this fraction is negligible.
      \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
      Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
      The color of the points corresponds to the tools used at each step of the pipeline (DC, TA, OP and NI).
      The points on the PCA plot can be grouped based on the TA step, but at the NI step only some methods create high variability in the inferred networks.
      The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot, but all the networks with filtering at the \ac{op} step are similar to each other.
      This further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.
    }
    \label{fig:figure6}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=1.0\linewidth]{figure7.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Comparison of networks generated from autistic and control samples using the \ac{micone} pipeline}.
      The networks for the control (left) and autistic (right) samples in the fecal microbiome transplant dataset were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
      There are 20 unique links in the network for control samples, 12 unique links in the network for autistic subjects and 7 edges in common between both networks.
    }
    \label{fig:figure7}
  \end{figure}

@@

    % TODO: Verify this about t-SNE
    % \begin{figure}[H]
    %   \centering
    %   \includegraphics[width=1.0\linewidth]{figure_s1.pdf}
    % \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The t-SNE plot of all the inferred networks clusters the networks based on the taxonomy reference database used}.
          Each point on the t-SNE plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
          The points are colored by the tools and parameters used in \ac{dc} step (A), \ac{ta} step (B), \ac{op} step (C) and \ac{ni} step (D).
          The separation of the points based on taxonomy reference database shows that the points might cluster based on reference database in high-dimensional space.
        }
      \label{fig:figure_s1}
    \end{figure}
    % \FloatBarrier
    % \newpage

>>>>

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=0.74\linewidth]{figure1.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The workflow of the \ac{micone} pipeline}.
      The steps of the workflow can be broken down into five major groups: \textbf{(SP)} \textbf{S}equence \textbf{P}rocessig, \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU and ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
      Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
      The text along the arrows describes the data that is being passed from one step to another.
      The final output of the pipeline is the consensus network generated from the inferred co-occurrence networks.
      For details on each process and the outputs, see Methods.
    }
    \label{fig:figure1}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure2.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The representative sequences generated by the different DC methods differ in their identification of sequences that are low in abundance.}
      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant (FMT) dataset~\cite{Kang2017}.
      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best-performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
      The data used for the analysis in \textbf{(C, D)} were the mock datasets from mockrobiota~\cite{Bokulich2016}.
    }
    \label{fig:figure2}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure3.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Taxonomic reference databases vary in terms of their taxonomy assignments beyond the Order level.}
      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
      This result illustrates how the same sequences are assigned to different Genus under different databases.
      A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
      The number of assigned genera for each database are displayed at the top of each column.
      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
      The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the FMT dataset.
      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
      The \ac{gg} database and the Naive Bayes classifier are chosen as the defaults for the TA step of \ac{micone}.
      The datasets used for the analysis in \textbf{(C)} were the mock datasets from mockrobiota.
    }
    \label{fig:figure3}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure4.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
      \textbf{(A)} The nine different networks (excluding \ac{mldm}) generated by the different network inference methods are very dissimilar.
      The green links are positive associations and the orange links represent negative associations.
      A threshold of 0.3 was set for the correlation-based methods (sparcc, propr, spearman and pearson) and a threshold of 0.01 was set for the direct association methods (flashweave, spieceasi, cozine, harmonies, and spring).
      The correlation-based methods in general produce graph with higher edge-densities.
      \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large proportion of common nodes involved in connections (33 out of 68).
      Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
      The data used in this analysis were the healthy stool samples from the FMT dataset.
      \ac{mldm} is not shown in the comparisons because the algorithm failed to converge for the particular network combination.
    }
    \label{fig:figure4}
  \end{figure}

@@

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure6.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
      \caption{
      \textbf{The choice of reference database contributes has the largest impact on network variance}.
      \textbf{(A)} The percentage of variance in the networks contributed by the DC, CC (chimera checking), TA, OP and NI steps of the pipeline calculated using ANOVA on the linear model, see the Methods section for details.
      The data used for the analysis were the healthy stool samples from the FMT dataset.
      A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the analysis.
      The taxonomy database contributes most to the variance between the networks (65.4\%) followed by the filtering of the counts matrix (26.8\%) in the OP step.
    The variation due to the DC and CC steps are much smaller in comparison (0.648\% and 0.003\% respectively).
      The fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time, however, this can be ignored as this fraction is negligible.
      \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
      Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
      The color of the points corresponds to the tools used at each step of the pipeline (DC, TA, OP and NI).
      The points on the PCA plot can be grouped based on the TA step, but at the NI step only some methods create high variability in the inferred networks.
      The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot, but all the networks with filtering at the \ac{op} step are similar to each other.
      This further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.
    }
    \label{fig:figure6}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=1.0\linewidth]{figure7.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Comparison of networks generated from autistic and control samples using the \ac{micone} pipeline}.
      The networks for the control (left) and autistic (right) samples in the FMT dataset were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
      There are 20 unique links in the network for control samples, 12 unique links in the network for autistic subjects and 7 edges in common between both networks.
    }
    \label{fig:figure7}
  \end{figure}

@@

    % \begin{figure}[H]
    %   \centering
    %   \includegraphics[width=1.0\linewidth]{figure_s1.pdf}
    % \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The t-SNE plot of all the inferred networks clusters the networks based on the taxonomy reference database used}.
          Each point on the t-SNE plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
          The points are colored by the tools and parameters used in \ac{dc} step (A), \ac{ta} step (B), \ac{op} step (C) and \ac{ni} step (D).
          The separation of the points based on taxonomy reference database shows that the points cluster based on reference database in high-dimensional space.
        }
      \label{fig:figure_s1}
    \end{figure}
    % \FloatBarrier
    % \newpage
